Understanding models for spatial structure in small-area estimation

### 2022-08-03 src/docs_paper/paper.Rmd
Update title

---
title: Modelling spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Harrison Zhu
    affiliation: A
  - name: Yidan Xu
    affiliation: C
  - name: Samir Bhatt
    affiliation: D
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Statistics, University of Michigan
  - code: D
    address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. Finally, we compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**

>>>>

---
title: Understanding models for spatial structure in small-area estimation
author:
  - name: Adam Howes
    email: ath19@ic.ac.uk
    affiliation: A
    corresponding: ath19@ic.ac.uk
  - name: Harrison Zhu
    affiliation: A
  - name: Yidan Xu
    affiliation: C
  - name: Samir Bhatt
    affiliation: D
  - name: Jeffrey W. Eaton
    affiliation: B
  - name: Seth R. Flaxman
    affiliation: E
address:
  - code: A
    address: Department of Mathematics, Imperial College London
  - code: B
    address: MRC Centre for Global Infectious Disease Analysis, Imperial College London
  - code: C
    address: Department of Statistics, University of Michigan
  - code: D
    address: Department of Public Health, University of Copenhagen
  - code: E
    address: Department of Computer Science, University of Oxford
abstract: |
  |  **Background:** Spatial correlation between small-areas is typically accounted for within a generalised linear mixed model setting using spatial random effects. However, commonly used models for spatial random effects based upon adjacency relations, like the Besag model, make unrealistic and unsatisfying spatial assumptions, which seem particularly dubious for irregular geometries. Furthermore, they do not take into account that across many settings areal data is generated by aggregating continuous point data. Defining spatial random effects which instead correspond to aggregated Gaussian processes gives a more intuitively convincing correlation structure, and allows tools from kernel methods to be used in models for areal data.
  |
  |  **Methods:** In a comprehensive simulation study, we used a proper scoring rule to evaluate the performance of different spatial random effect specifications. We performed these simulations under model misspecification and across a range of vignette and realistic geometries with different regularities. Finally, we compared model performance in estimating district-level HIV prevalence from PHIA household survey data across four countries in sub-Saharan Africa, using cross-validation and spatial cross-validation.
  |
  |  **Results:**
  |
  |  **Conclusions:**
